Distinguishing multiple-merger from Kingman coalescence using two-site frequency spectra

### 2018-08-29 manuscript/manuscript.tex
Add missing citations.

One of the most widely used models is the Kingman coalescent~\autocite{Kingman1982a, Kingman1982b, Kingman1982c, Hudson1983, Tajima1983}.
The Kingman coalescent is a stochastic process that generates gene genealogies: trees representing the patterns of shared ancestry of sampled individuals.
Inference methods use these genealogies as latent variables linking demographic parameters to genetic data~\autocite{RosenbergNordborg2002}.
The Kingman coalescent has a number of convenient properties that allow for both analytical calculations (e.g.,~\cite{Tajima1989}) and efficient stochastic simulations (e.g.,~\cite{Hudson2002}): tree topologies are independent of waiting times; waiting times are generated by a Markov process; and neutral mutations are modeled as a Poisson process conditionally independent of the tree.
Moreover, the model can be extended to study a variety of biological phenomena including recombination, population structure, and variation in sex ratios or ploidy (see generally~\cite{Wakeley2009}).

An important application of the Kingman coalescent is inferring historical population sizes from genetic data~\autocite{SchraiberAkey2015}.
In its simplest form, the model has a single parameter, the coalescent rate, which determines the branch lengths of genealogies~\autocite{Kingman1982a}.
Under a variety of conditions, the coalescent rate is inversely proportional to the population size~\autocite{Kingman1982b}.
Accordingly, a growing or shrinking population may be modeled by a time-varying rate~\autocite{GriffithsTavare1994, GriffithsTavare1998}.
Patterns of genetic diversity depend on the ratio of the coalescent rate to other evolutionary rate parameters.
For example, the \emph{site frequency spectrum} (SFS)---the number of mutations segregating at different frequencies in a sample---is determined by the ratio of the mutation rate to the (time-varying) coalescent rate.
Kingman coalescent--based inference methods solve the inverse problem of determining the population size history that best explains particular features of the data, such as the SFS (e.g., \cite{BhaskarEtAl2015}) or variations in heterozygocity along a chromosome (e.g., \cite{LiDurbin2011}).

@@

An important extension of the Kingman coalescent is a family of models known as \textit{multiple merger coalescents} \autocite{Pitman1999, Sagitov1999, DonnellyKurtz1999,} (reviewed in \cite{Eldon2016}), which arise in a variety of contexts both with and without selection.
Whereas in the Kingman coalescent lineages may coalesce only pairwise, multiple merger coalescents permit more than two lineages to coalesce in a single event.
The even more general class of simultaneous multiple merger coalescents \autocite{Schweinsberg2000, MohleSagitov2001, Sagitov2003} permits more than one distinct multiple merger event at the same time.
These models are relevant for species with
``sweepstakes'' reproductive events \autocite{EldonWakeley2006, SargsyanWakeley2008},
fat-tailed offspring number distributions \autocite{Schweinsberg2003},
recurring selective sweeps at linked sites \autocite{DurrettSchweinsberg2005, CoopRalph2012},
rapid adaptation \autocite{NeherHallatscheck2013, DesaiEtAl},
and purifying selection at sufficiently dense sites \autocite{Seger, Good, Nicholaisen}.

@@

One approach to identifying multiple mergers in genomic data is to use the SFS as a summary statistic.
\cite{BirknerEtAl2013, BlathEtAl2016, SpenceEtAl2016} derived methods for computing the expected site frequency spectrum of (simultaneous-)multiple merger coalescents.
\cite{EldonEtAl2015} showed that it is possible to distinguish between a multiple merger coalescent of the beta family and the Kingman coalescent with exponential growth using the SFS.
\cite{RodelspergerEtAl2014} detected widespread linked selection in the nematode \textit{Pristionchus pacificus} by demonstrating that the site frequency spectrum is non-monotonic, a signature of multiple mergers \autocite{NeherHallatscheck2013, BirknerEtAl2013}.

@@

In contrast, the 2-SFS is a statistic of \textit{pairs} of sites.
We define the 2-SFS,
$\left\{ \xi_{ij}(d) : d > 0; 1 \leq i, j < n\right\}$,
as the fraction of pairs of sites separated by $d$ bases for which there is a mutation with derived allele count $i$ at one site and a second mutation with derived allele count $j$ at the other site.
(Note that $\xi_{ij}(d) = \xi_{ji}(d)$ by symmetry.)
This object has been studied for non-recombining sites by \cite{FerrettiEtAl2018} in a neutral model and \cite{Xie2011} in a model with selection.
We define the folded 2-SFS, $\eta_{ij}(d)$, by analogy to the folded SFS,  categorizing pairs of sites by their minor allele frequencies. (For non-recombining sites, the 2-SFS is independent of the distance and so we will suppress the $d$ in our notation.)

@@

\cite{BirknerEtAl2013} extended Fu's calculation to a family of multiple merger coalescents called beta coalescents.
This one-parameter family interpolates between the Kingman coalescent and the Bolthausen-Sznitman coalescent as the parameter, $\alpha$, varies from 2 to 1.
Beta coalescents arise in models with fat-tailed offspring distributions \autocite{Schweinsberg2003}, and the Bolthausen-Sznitman coalescent is the limiting distribution of genealogies in rapidly adapting populations \autocite{NeherHallatscheck2013}.
Like Fu, Birkner et al.\ were primarily concerned with computing the sample variance of SFS-based summary statistics such as Tajima's D \autocite{Tajima1989}.
As a result, they were mostly interested in the diagonal terms of the SFS covariance matrix, which dominate that calculation.
However, Figures 5 and 6 of \cite{BirknerEtAl2013} show positive correlations between $\xi_i$ and $\xi_j$ for $j \not\in \{i, n-i\}$.
Thus, unlike the standard Kingman coalescent, the beta coalescent can generate positive associations between mutations with different minor allele counts.

@@

We ran coalescent simulations using a custom version of \texttt{msprime} \autocite{KelleherEtAl2014} capable of multiple mergers, based on modifications made by Joe Zhu [GITHUB].
This code, together with \texttt{python} wrapper scripts and utility functions to run simulations and calculate the 2-SFS and $\fpmi$ from the \texttt{msprime} output is available [GITHUB].

@@

We simulated a model of recurring selective sweeps using the software \texttt{SLiM} \autocite{MesserEtAl201?}.
In all of these simulations, we simulated a population of 500 diploids for $10^4$ generations.
Each haploid genome consisted of a single genomic element $L=10^8$ basepairs long with recombination rate per basepair $r = 10^{-8}$ and overall mutation rate $\mu = 10^{-7}$.
We simulated two types of mutations: neutral mutations and beneficial mutations with additive effects and selection coefficient $s=0.1$.
With these parameters, $2Ns = 100$, so beneficial mutations are strongly selected and will sweep in $T_{\text{sweep}}\sim s^{-1} \log Ns \approx 50$ generations.
Such sweeps will effect a region of the chromosome $d_{\text{sweep}} \sim (r T_{\text{sweep}})^{-1} \approx 2 \times 10^6$ basepairs long.
Thus $d_{\text{sweep}} \ll L$, which will minimize edge effects of simulating a finite chromosome.

@@

In addition to the average SFS used to fit the model, we computed the average 2-SFS for pairs of sites at distances distances between 3~bp and 5~Kb.
Because we are using four-fold degenerate sites, we only computed the 2-SFS for distances that are multiples of 3.
For comparison between data and simulations, we scale the distances in basepairs by a critical distance $d_c = (r\E{T_2})^{-1}$.
We used a genome-wide recombination rate of $r = 2 \times 10^{-8}$ per basepair per generation \autocite{ComeronEtAl2012}.
We estimated $\E{T_2}$ by $\Pi / 2 \mu$.
For $\Pi$, we used the average pairwise diversity at fourfold degenerate sites in the central region of each chromosome arm.
For $\mu$, we used a genome-wide mutation rate of $3 \times 10^{-9}$ per basepair per generation \autocite{KeightlyEtAl2014}.
These estimates are not precise, but only serve to scale genetic distances to the correct order of magnitude.

@@

We compared the $\fpmi$ of the Kingman coalescent with and without population growth to the $\fpmi$ of the beta coalescent, for pairs of sites without recombination.
We are interested in whether $\fpmi$ can distinguish beta from Kingman coalescent models that produce similar distortions in the SFS.
To this end, we computed a version of Tajima's D \autocite{Tajima1989} normalized to be invariant under changes in the mean coalescent rate: $D = (\Pi - \hat{\theta}_W) / \Pi$, where $\hat{\theta}_W$ is Watterson's theta, a linear function of the SFS \autocite{Watterson19??}.
Negative values of $D$ indicate an overabundance of low-frequency mutations relative to the time-homogeneous Kingman expectation.
All results are computed numerically using the results of \cite{Fu1995}, \cite{ZivkovicWiehe2008}, and \cite{BirknerEtAl2013} (see Methods).

@@

There is one other salient feature of \fig{fig:hilo_vs_d}: $\hilopmi$ does not decay to zero at long distances in the beta coalescent.
This is related to the results of \cite{EldonWakeley20??}, who showed that a model with ``jackpot'' reproductive events can generate infinite-range linkage disequilibrium.
They showed that different scalings of rates of mutation, pairwise coalescence, multiple merger coalescence, and recombination lead to different behaviors of diversity and linkage disequilibrium.
Our implementation of the beta coalescent model with recombination corresponds to a particular scaling limit where recombination does not have time to decorrelate trees during multiple mergers events, even at infinite genetic distances.
We do not expect this behavior to be universal in multiple mergers coalescents with recombination.

@@

Various authors have shown that natural selection can generate multiple merger coalescents at linked neutral sites (e.g., \cite{DurrettSchweinsberg2005, CoopRalph, NeherHallatscheck2013, DesaiEtAl, Seger}).
However, our simulations of the beta coalescent with recombination are of an explicitly neutral model.
Thus, they are at best an approximation to the selective models cited above.
It is therefore important to verify that selection at linked sites can, in fact, generate the sorts of signals in $\fpmi$ that we have detected in the beta coalescent.

To test this proposition, we performed forward-time simulations of recurring selective sweeps using the software \texttt{SLiM} \autocite{MesserEtAl201?}.
In these simulations, we simulated individual chromosomes with homogenous recombination and two types of mutations: neutral and beneficial with fixed selection coefficient $s$.
Both types of mutations occurred at random, uniformly distributed across the length of the chromosome.
By varying the recombination rate, mutation rates, population size, and selection coefficient, we were able to vary the rate of selective sweeps linked to neutral sites.

@@

In this section, we illustrate the procedure just outlined by using genomic diversity data from the \textit{Drosophila melanogaster} DGPG3 panel \cite{LackEtAl2015}.
The DPGP3 data consists of haploid consensus sequences from $\sim$ 200 wild-caught flies from a Zambian population known to be relatively free of cosmopolitan admixture.
Recently, several groups have used the DPGP3 data to estimate the population-size history of \textit{D. melanogaster} \autocite{TerhorstEtAl2017,RagsdaleGutenkunst2017}.
On the other hand, it is widely believed that the genetic diversity of \textit{Drosophila} is strongly shaped by natural selection (e.g., \cite{Elyashiv??201?, GarudPetrov2016, others?})
Thus, this data is a good candidate for demonstrating the utility of $\fpmi$ for assessing coalescent model fit.

@@

\section*{Discussion}
[DEAL WITH THIS: In fact, for large non-recombining loci, it is sometimes possible to reconstruct the gene genealogy explicitly and calculate tree imbalance statistics (see e.g., \cite{Seger2010}). HERE WE DONT NEED A TREE]

@@

We focus here on demographic inference methods that explicitly use the Kingman coalescent model for calculations.
However, another popular class of methods are based on forward-time models, such as the Wright-Fisher model (e.g., \cite{GutenkunstEtAl2009}, \cite{dical}).
We believe that $\fpmi$ is useful for model-checking with these methods as well.
This is because forward-time neutral models each have an associated dual coalescent model \cite{etheridge}, which can be used to compute the expected $\fpmi$ as outlined in this paper.
Furthermore, non-neutral models also generate predictions about the 2-SFS and thus $\fpmi$.
In principle, any fitted model that allows calculation or simulation of the 2-SFS may be checked using $\fpmi$.
However, the suitability of $\fpmi$ for discriminating among arbitrary families of models is unknown.

In this paper, we have outlined a graphical model-checking procedure in the spirit of \cite{Anscombe} and \cite{GelmanBDA}.
One could extend our work to implement a formal hypothesis-testing framework for rejecting the best-fit Kingman coalescent model according to some inference method.
Numerical calculations would require higher moments of the branch-length distribution, which may be computationally intractable for large samples.
On the other hand, it would be straightforward to develop a bootstrap-style procedure, using coalescent simulations to estimate the variance in test statistics under the fit model.
In any case, we believe that it would be useful for simulation packages such as \texttt{msprime} and population genetics analysis libraries to include standard functions for computing the 2-SFS, $\fpmi$, and $\hilopmi$.
We provide such functions in the [GITHUB] repository associated with this paper.

@@

\section*{Supplementary Methods}
[MOVE STUFF HERE IF NECESSARY]

@@

@@

OLD TEXT:
% Tests of the standard Kingman coalescent based on the SFS originated with \cite{Tajima1989}, \cite{FuLi1993}, and \cite{SimonsenEtAl1995} (see \cite{Achaz2009} for a unifying framework).

>>>>

One of the most widely used models is the Kingman coalescent~\autocite{Kingman1982, Kingman1982a, Hudson1983, Tajima1983}.
The Kingman coalescent is a stochastic process that generates gene genealogies: trees representing the patterns of shared ancestry of sampled individuals.
Inference methods use these genealogies as latent variables linking demographic parameters to genetic data~\autocite{RosenbergNordborg2002}.
The Kingman coalescent has a number of convenient properties that allow for both analytical calculations (e.g.,~\cite{Tajima1989}) and efficient stochastic simulations (e.g.,~\cite{Hudson2002}): tree topologies are independent of waiting times; waiting times are generated by a Markov process; and neutral mutations are modeled as a Poisson process conditionally independent of the tree.
Moreover, the model can be extended to study a variety of biological phenomena including recombination, population structure, and variation in sex ratios or ploidy (see generally~\cite{Wakeley2009}).

An important application of the Kingman coalescent is inferring historical population sizes from genetic data~\autocite{SchraiberAkey2015}.
In its simplest form, the model has a single parameter, the coalescent rate, which determines the branch lengths of genealogies~\autocite{Kingman1982}.
Under a variety of conditions, the coalescent rate is inversely proportional to the population size~\autocite{Kingman1982a}.
Accordingly, a growing or shrinking population may be modeled by a time-varying rate~\autocite{GriffithsTavare1994, GriffithsTavare1998}.
Patterns of genetic diversity depend on the ratio of the coalescent rate to other evolutionary rate parameters.
For example, the \emph{site frequency spectrum} (SFS)---the number of mutations segregating at different frequencies in a sample---is determined by the ratio of the mutation rate to the (time-varying) coalescent rate.
Kingman coalescent--based inference methods solve the inverse problem of determining the population size history that best explains particular features of the data, such as the SFS (e.g., \cite{BhaskarEtAl2015}) or variations in heterozygocity along a chromosome (e.g., \cite{LiDurbin2011}).

@@

An important extension of the Kingman coalescent is a family of models known as \textit{multiple merger coalescents} \autocite{Pitman1999, Sagitov1999, DonnellyKurtz1999,} (reviewed in \cite{Eldon2016}), which arise in a variety of contexts both with and without selection.
Whereas in the Kingman coalescent lineages may coalesce only pairwise, multiple merger coalescents permit more than two lineages to coalesce in a single event.
The even more general class of simultaneous multiple merger coalescents \autocite{Schweinsberg2000, MohleSagitov2001, Sagitov2003} permits more than one distinct multiple merger event at the same time.
These models are relevant for species with
``sweepstakes'' reproductive events \autocite{EldonWakeley2006, SargsyanWakeley2008},
fat-tailed offspring number distributions \autocite{Schweinsberg2003},
recurring selective sweeps at linked sites \autocite{DurrettSchweinsberg2005, CoopRalph2012},
rapid adaptation \autocite{NeherHallatschek2013, DesaiEtAl2013},
and purifying selection at sufficiently dense sites \autocite{SegerEtAl2010, NicolaisenDesai2012}.

@@

One approach to identifying multiple mergers in genomic data is to use the SFS as a summary statistic.
\cite{BirknerEtAl2013, BlathEtAl2016, SpenceEtAl2016} derived methods for computing the expected site frequency spectrum of (simultaneous-)multiple merger coalescents.
\cite{EldonEtAl2015} showed that it is possible to distinguish between a multiple merger coalescent of the beta family and the Kingman coalescent with exponential growth using the SFS.
\cite{RodelspergerEtAl2014} detected widespread linked selection in the nematode \textit{Pristionchus pacificus} by demonstrating that the site frequency spectrum is non-monotonic, a signature of multiple mergers \autocite{NeherHallatschek2013, BirknerEtAl2013}.

@@

In contrast, the 2-SFS is a statistic of \textit{pairs} of sites.
We define the 2-SFS,
$\left\{ \xi_{ij}(d) : d > 0; 1 \leq i, j < n\right\}$,
as the fraction of pairs of sites separated by $d$ bases for which there is a mutation with derived allele count $i$ at one site and a second mutation with derived allele count $j$ at the other site.
(Note that $\xi_{ij}(d) = \xi_{ji}(d)$ by symmetry.)
This object has been studied for non-recombining sites by \cite{FerrettiEtAl2018} in a neutral model and \cite{Xie2011a} in a model with selection.
We define the folded 2-SFS, $\eta_{ij}(d)$, by analogy to the folded SFS,  categorizing pairs of sites by their minor allele frequencies. (For non-recombining sites, the 2-SFS is independent of the distance and so we will suppress the $d$ in our notation.)

@@

\cite{BirknerEtAl2013} extended Fu's calculation to a family of multiple merger coalescents called beta coalescents.
This one-parameter family interpolates between the Kingman coalescent and the Bolthausen-Sznitman coalescent as the parameter, $\alpha$, varies from 2 to 1.
Beta coalescents arise in models with fat-tailed offspring distributions \autocite{Schweinsberg2003, SteinruckenEtAl2013}, and the Bolthausen-Sznitman coalescent is the limiting distribution of genealogies in rapidly adapting populations \autocite{NeherHallatschek2013}.
Like Fu, Birkner et al.\ were primarily concerned with computing the sample variance of SFS-based summary statistics such as Tajima's D \autocite{Tajima1989}.
As a result, they were mostly interested in the diagonal terms of the SFS covariance matrix, which dominate that calculation.
However, Figures 5 and 6 of \cite{BirknerEtAl2013} show positive correlations between $\xi_i$ and $\xi_j$ for $j \not\in \{i, n-i\}$.
Thus, unlike the standard Kingman coalescent, the beta coalescent can generate positive associations between mutations with different minor allele counts.

@@

We ran coalescent simulations using a custom version of \texttt{msprime} \autocite{KelleherEtAl2016} capable of multiple mergers, based on modifications made by Joe Zhu [GITHUB].
This code, together with \texttt{python} wrapper scripts and utility functions to run simulations and calculate the 2-SFS and $\fpmi$ from the \texttt{msprime} output is available [GITHUB].

@@

We simulated a model of recurring selective sweeps using the software \texttt{SLiM} \autocite{Messer2013}.
In all of these simulations, we simulated a population of 500 diploids for $10^4$ generations.
Each haploid genome consisted of a single genomic element $L=10^8$ basepairs long with recombination rate per basepair $r = 10^{-8}$ and overall mutation rate $\mu = 10^{-7}$.
We simulated two types of mutations: neutral mutations and beneficial mutations with additive effects and selection coefficient $s=0.1$.
With these parameters, $2Ns = 100$, so beneficial mutations are strongly selected and will sweep in $T_{\text{sweep}}\sim s^{-1} \log Ns \approx 50$ generations.
Such sweeps will effect a region of the chromosome $d_{\text{sweep}} \sim (r T_{\text{sweep}})^{-1} \approx 2 \times 10^6$ basepairs long.
Thus $d_{\text{sweep}} \ll L$, which will minimize edge effects of simulating a finite chromosome.

@@

In addition to the average SFS used to fit the model, we computed the average 2-SFS for pairs of sites at distances distances between 3~bp and 5~Kb.
Because we are using four-fold degenerate sites, we only computed the 2-SFS for distances that are multiples of 3.
For comparison between data and simulations, we scale the distances in basepairs by a critical distance $d_c = (r\E{T_2})^{-1}$.
We used a genome-wide recombination rate of $r = 2 \times 10^{-8}$ per basepair per generation \autocite{ComeronEtAl2012}.
We estimated $\E{T_2}$ by $\Pi / 2 \mu$.
For $\Pi$, we used the average pairwise diversity at fourfold degenerate sites in the central region of each chromosome arm.
For $\mu$, we used a genome-wide mutation rate of $3 \times 10^{-9}$ per basepair per generation \autocite{KeightleyEtAl2014}.
These estimates are not precise, but only serve to scale genetic distances to the correct order of magnitude.

@@

We compared the $\fpmi$ of the Kingman coalescent with and without population growth to the $\fpmi$ of the beta coalescent, for pairs of sites without recombination.
We are interested in whether $\fpmi$ can distinguish beta from Kingman coalescent models that produce similar distortions in the SFS.
To this end, we computed a version of Tajima's D \autocite{Tajima1989} normalized to be invariant under changes in the mean coalescent rate: $D = (\Pi - \hat{\theta}_W) / \Pi$, where $\hat{\theta}_W$ is Watterson's theta, a linear function of the SFS \autocite{Watterson1975}.
Negative values of $D$ indicate an overabundance of low-frequency mutations relative to the time-homogeneous Kingman expectation.
All results are computed numerically using the results of \cite{Fu1995}, \cite{ZivkovicWiehe2008}, and \cite{BirknerEtAl2013} (see Methods).

@@

There is one other salient feature of \fig{fig:hilo_vs_d}: $\hilopmi$ does not decay to zero at long distances in the beta coalescent.
This is related to the results of \cite{EldonWakeley2006}, who showed that a model with ``jackpot'' reproductive events can generate infinite-range linkage disequilibrium.
They showed that different scalings of rates of mutation, pairwise coalescence, multiple merger coalescence, and recombination lead to different behaviors of diversity and linkage disequilibrium.
Our implementation of the beta coalescent model with recombination corresponds to a particular scaling limit where recombination does not have time to decorrelate trees during multiple mergers events, even at infinite genetic distances.
We do not expect this behavior to be universal in multiple mergers coalescents with recombination.

@@

Various authors have shown that natural selection can generate multiple merger coalescents at linked neutral sites (e.g., \cite{DurrettSchweinsberg2005, CoopRalph2012, NeherHallatschek2013, DesaiEtAl2013, SegerEtAl2010}).
However, our simulations of the beta coalescent with recombination are of an explicitly neutral model.
Thus, they are at best an approximation to the selective models cited above.
It is therefore important to verify that selection at linked sites can, in fact, generate the sorts of signals in $\fpmi$ that we have detected in the beta coalescent.

To test this proposition, we performed forward-time simulations of recurring selective sweeps using the software \texttt{SLiM} \autocite{Messer2013}.
In these simulations, we simulated individual chromosomes with homogenous recombination and two types of mutations: neutral and beneficial with fixed selection coefficient $s$.
Both types of mutations occurred at random, uniformly distributed across the length of the chromosome.
By varying the recombination rate, mutation rates, population size, and selection coefficient, we were able to vary the rate of selective sweeps linked to neutral sites.

@@

In this section, we illustrate the procedure just outlined by using genomic diversity data from the \textit{Drosophila melanogaster} DGPG3 panel \cite{LackEtAl2015}.
The DPGP3 data consists of haploid consensus sequences from $\sim$ 200 wild-caught flies from a Zambian population known to be relatively free of cosmopolitan admixture.
Recently, several groups have used the DPGP3 data to estimate the population-size history of \textit{D. melanogaster} \autocite{TerhorstEtAl2017,RagsdaleGutenkunst2017}.
On the other hand, it is widely believed that the genetic diversity of \textit{Drosophila} is strongly shaped by natural selection (e.g., \cite{ElyashivEtAl2016, GarudPetrov2016})
Thus, this data is a good candidate for demonstrating the utility of $\fpmi$ for assessing coalescent model fit.

@@

\section*{Discussion}

@@

We focus here on demographic inference methods that explicitly use the Kingman coalescent model for calculations.
However, another popular class of methods are based on forward-time models, such as the Wright-Fisher model (e.g., \cite{GutenkunstEtAl2009}, \cite{SheehanEtAl2013}).
We believe that $\fpmi$ is useful for model-checking with these methods as well.
This is because forward-time neutral models each have an associated dual coalescent model \cite{Etheridge2011e}, which can be used to compute the expected $\fpmi$ as outlined in this paper.
Furthermore, non-neutral models also generate predictions about the 2-SFS and thus $\fpmi$.
In principle, any fitted model that allows calculation or simulation of the 2-SFS may be checked using $\fpmi$.
However, the suitability of $\fpmi$ for discriminating among arbitrary families of models is unknown.

In this paper, we have outlined a graphical model-checking procedure in the spirit of \cite{Anscombe1973}.
One could extend our work to implement a formal hypothesis-testing framework for rejecting the best-fit Kingman coalescent model according to some inference method.
Numerical calculations would require higher moments of the branch-length distribution, which may be computationally intractable for large samples.
On the other hand, it would be straightforward to develop a bootstrap-style procedure, using coalescent simulations to estimate the variance in test statistics under the fit model.
In any case, we believe that it would be useful for simulation packages such as \texttt{msprime} and population genetics analysis libraries to include standard functions for computing the 2-SFS, $\fpmi$, and $\hilopmi$.
We provide such functions in the [GITHUB] repository associated with this paper.

@@

@@

We also don't mention single locus methods.

@@

OLD TEXT:

% Tests of the standard Kingman coalescent based on the SFS originated with \cite{Tajima1989}, \cite{FuLi1993}, and \cite{SimonsenEtAl1995} (see \cite{Achaz2009} for a unifying framework).
